---
title: "Allosteric Cooperativity in Tetramer Mixture"
output:
  html_document: default
  html_notebook: default
---

## Posits
1. We consider a mixture of two components A (active) and I (inactive).
2. The components form exclusively tetramers.
3. The composition of tetramers is purely stochastic.
4. The mixture is defined by the ratio of the concentrations of A and I.


## Configurations
We have 5 possible configurations, not counting any permutations:
AAAA, AAAI, AAII, AIII, IIII


## Nomenclature
* concentration a = c(A)
* concentration i = c(I)
* concentration ratio *r*; ***[eq1]***
```{r, eval = FALSE}
r = a / i
```
* total concentration *C*; ***[eq2]***
```{r, eval = FALSE}
C = a + i
```

## Concentrations
### Concentration in terms of *r*; ***[eq3]***
```{r, eval = FALSE}
r = 10; # in practice (below) a sequence from 0 to 10
a = r * i;
i = a / r;
```

### Concentration *a*; ***[eq4]*** in ***[eq2]***
```{r, eval = FALSE}
a + (a / r) = C;
a (1 + (1 / r)) = C;
a = C / (1 + (1 / r));
```

### Concentration *i*; ***[eq3]*** in ***[eq2]***
```{r, eval = FALSE}
r * i + i = C;
i (1 + r) = C;
i = C / (1 + r);
```

## Constants: Total concentration and allosteric coefficient
```{r}
C = 1;
AC = 4;
```

## Models for enzymatic activities of tetramers

1. Non-cooperative (linear) model
```{r}
aaaa_ea = aaai_ea = aaii_ea = aiii_ea = iiii_ea = 1;
```
2. Pair model: number of possible active pairs
```{r}
aaaa_ea = 3^AC;
aaai_ea = 2^AC;
aaii_ea = 1^AC;
aiii_ea = (1/4);
iiii_ea = (0/4);
```

3. Cooperative (non-linear) model
```{r}
aaaa_ea = (4/4)^AC;
aaai_ea = (3/4)^AC;
aaii_ea = (2/4);
aiii_ea = (1/4);
iiii_ea = (0/4);
```

## Evaluate enzmatic activity (ea) of the tetramer mixture for a range of A/I concentration ratios
```{r}
rs = seq(0.1, 10, by = 0.01);

result.ea = lapply(rs, function(r) {
    a = C / (1 + 1 / r);
    i = C / (1 + r);
    sum(a + i);

    #_______________________________________________________________________________
    ## tetramer concentrations
    aaaa = a^4;
    aaai = 4 * a^3 * i;
    aaii = 6 * a^2 * i^2;
    aiii = 4 * a * i^3;
    iiii = i^4;
    stopifnot(isTRUE(all.equal(sum(aaaa + aaai + aaii + aiii + iiii), 1)));

    #_______________________________________________________________________________
    ## total enzymatic activity of tetramer mixture
    total_ea =  aaaa * aaaa_ea +
                aaai * aaai_ea +
                aaii * aaii_ea +
                aiii * aiii_ea +
                iiii * iiii_ea;

    c(i, total_ea);
});

result.ea.df = do.call(rbind.data.frame, result.ea);
colnames(result.ea.df) = c("i", "total_ea");
plot(abs(result.ea.df[ ,1]), result.ea.df[ ,2], xlab = c("i"), ylab = c("enzymatic activity"));
```


